Digging Deeper Into the State of the Practice for Domain Specific Research Software

### 2022-02-18 StateOfPractice/Papers/ICCS-2022-SOP_MethodologyPlusExamples/DigDeep.tex
Revisions to Conclusions section, fixed communities to community's.  Done for now.

To address \rqref{RQ_CompareHQ2Popular} we need to compare the ranking by best
practices to the community's ranking.  Our best practices ranking comes from the
AHP ranking (Section~\ref{empiricalmeasures}).  We estimate the community's
ranking by repository stars and watches.  The comparison will provide insight on
whether best practices are rewarded by popularity.  However, inconsistencies
between the AHP ranking and the communities ranking are inevitable for the
following reasons: 
\begin{inparaenum}[i)]
	\item the overall quality ranking via AHP makes the unrealistic assumption
	of equal weighting between quality factors;
	\item stars are known to not be a particularly good measure of popularity,
	because of how people use stars and because young projects have less time to
	accumulate stars~\cite{Szulik2017}; 
	\item and, as for consumer products, there are more factors influencing
	popularity than just quality.
\end{inparaenum}

@@

We have outlined a methodology for assessing the state of the practice for any
given research software domain based on a list of about 30 representative
projects.  In addition to interviews, we use manual and automated inspection of
the artifacts in each project's repositories.  The repository data is collected
by filling in a 108 question measurement template, which requires installing the
software on a VM, running simple tests (like completing the getting started
instructions (if present)), and searching the code, documentation and test
files. Using AHP the projects are ranked for 9 individual qualities
(installability, correctness and verifiability, reliability, robustness,
usability, maintainability, reusability, surface understandability,
visibility/transparency) and for overall quality.  Perspective and insight is
shared with the user community via the following: i) comparing the ranking by
best practices against an estimate of the communities ranking of popularity; ii)
comparing artifacts, tools and processes to current research software
development guidelines; and, iii) exploring pain points via developer
interviews. Using our methodology, spreadsheet templates and AHP tool, we
estimate (based on our experience with using the process) the time to complete
an assessment for a given domain at 173 person hours.

>>>>

To address \rqref{RQ_CompareHQ2Popular} we need to compare the ranking by best
practices to the community's ranking.  Our best practices ranking comes from the
AHP ranking (Section~\ref{empiricalmeasures}).  We estimate the community's
ranking by repository stars and watches.  The comparison will provide insight on
whether best practices are rewarded by popularity.  However, inconsistencies
between the AHP ranking and the community's ranking are inevitable for the
following reasons: 
\begin{inparaenum}[i)]
	\item the overall quality ranking via AHP makes the unrealistic assumption
	of equal weighting between quality factors;
	\item stars are known to not be a particularly good measure of popularity,
	because of how people use stars and because young projects have less time to
	accumulate stars~\cite{Szulik2017}; 
	\item and, as for consumer products, there are more factors influencing
	popularity than just quality.
\end{inparaenum}

@@

We have outlined a methodology for assessing the state of the practice for any
given research software domain based on a list of about 30 representative
projects.  In addition to interviews, we use manual and automated inspection of
the artifacts in each project's repositories.  The repository data is collected
by filling in a 108 question measurement template, which requires installing the
software on a VM, running simple tests (like completing the getting started
instructions (if present)), and searching the code, documentation and test
files. Using AHP the projects are ranked for 9 individual qualities
(installability, correctness and verifiability, reliability, robustness,
usability, maintainability, reusability, surface understandability,
visibility/transparency) and for overall quality.  Perspective and insight is
shared with the user community via the following: i) comparing the ranking by
best practices against an estimate of the community's ranking of popularity; ii)
comparing artifacts, tools and processes to current research software
development guidelines; and, iii) exploring pain points via developer
interviews. Using our methodology, spreadsheet templates and AHP tool, we
estimate (based on our experience with using the process) the time to complete
an assessment for a given domain at 173 person hours.
